---
title: "Iris Dataset"
date: "2024-02-15"
output: html_document
---
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1 
```{r Q1, echo = FALSE, include=FALSE}
data(iris)
iris
help(iris)

colnames(iris)
rownames(iris)

iris [10,]
iris[,3]
iris$Petal.Length
subset_data = iris[10:20, 1:2]
subset_data
species_names = iris$Species
```

a.) The variables are: sepal length and width and petal length and width and species.
For the 10th sample, the Sepal length is 4.9, Sepal width is 3.1, Petal length is 1.5, Petal Width is 0.1, and the Species is Setosa.
```{r, echo = TRUE}
iris[10,]
```

b.) There are 150 samples in this dataset.
The different species that were measured are: Setosa, Versicolor, and Virginica.
```{r}
#num of samples
num_rows = nrow(iris)
num_rows
#num of species
unique(iris$Species)
```

# Question 2
```{r}
summary(iris)
sd_cols = apply(iris[,1:4], 2, sd)
sd_rows = apply(iris[,1:4], 1, sd)
boxplot(iris[, 1:4], main = "Boxplot of Iris Variables")
```
```{r Q2}
#standard deviation
stdev = sd(iris$Sepal.Length)
stdev

#5 num summary
sub_sum = summary(subset_data)
print(sub_sum)
sub_set = iris[10:20, "Sepal.Length"]
standard_deviation = sd(sub_set)
print(standard_deviation)
```
a.) There are 50 observations of each species.
b.) Sepal.Length and Sepal.Width have medians smaller than their means.
c.) The standard deviation of the sepal length measurements is 0.8280661.
d.) The standard deviation of the sepal length measurements of this subset is 0.4665151.


# Question 3
```{r Q3}
atomic_vec = c(1, 4, 3, 2, NA, 3.22, -44, 2, NA, 0, 22, 34)
```

a.) There are 8 positive numbers in the vector.
b.) There is 1 negative number in the vector,
c.) There is 1 zero in the vector.
d.) There are 2 NAs in the vector.
e.) There are 9 non-zero & NA numbers in the vector.
f.) The sum of the positive numbers is 71.22.
g.) The sum of the negative numbers is -44.
```{r}
#a
pos_nums = sum(atomic_vec > 0, na.rm = TRUE)
print(pos_nums)
#b
neg_nums = sum(atomic_vec < 0, na.rm = TRUE)
print(neg_nums)
#c
zero_nums = sum(atomic_vec == 0, na.rm = TRUE)
print(zero_nums)
#d
NA_nums = sum(is.na(atomic_vec))
print(NA_nums)
#e
e = sum(atomic_vec != 0 & !is.na(atomic_vec))
print(e)
#f
sums_pos = sum(atomic_vec[atomic_vec>0], na.rm = TRUE)
print(sums_pos)
#g
sums_neg = sum(atomic_vec[atomic_vec<0], na.rm = TRUE)
print(sums_neg)
```

# Question 4
a.)
```{r, include = FALSE, echo = FALSE}
#create vector
fib_vec = rep(0,1000)
fib_vec[1] = 1
fib_vec[2] = 1
for (i in 3:1000) {
  fib_vec[i] = fib_vec[i-1] + fib_vec[i-2]
} 
```
b.) First 8 entries:  1,  1,  2,  3,  5, 8, 13, 21.
    Last 8 entries:  1.497069e+207, 2.422308e+207, 3.919377e+207,6.341685e+207, 1.026106e+208, 1.660275e+208, 2.686381e+208, 4.346656e+208.
```{r}
first = fib_vec[1:8]
print(first)
last = tail(fib_vec,8)
print(last)

```

c.)
```{r}
z_n = fib_vec[-length(fib_vec)] / fib_vec[-1]
z_n999 = z_n[1:999]
head(z_n999,10)

```
d.) 
```{r}
plot(z_n[1:10])
abline(h = (1 + sqrt(5))/2)
```

e.) Initially there is a big change but the values become more stable and similar as the n-value increases.

f.) This code is not meant for vectors. The syntax for vectors would need brackets.


# Question 5

```{r}
#matrix 1
fib_matrix1 = matrix(fib_vec, nrow = 100, ncol = 10, byrow = TRUE)
#print(fib_matrix1)

#matrix 2
fib_matrix2 = t(fib_matrix1)
#print(fib_matrix2)
```

```{r}
#array
fib_vector = numeric(100)
fib_vector[1] = 1
fib_vector[2] = 1
for (i in 3:100){
  fib_vector[i] = fib_vector[i-1] + fib_vector[i-2]
}

fib_matrices = array(fib_vector, dim = c(10,10,10))
fib_stacked = apply(fib_matrices, MARGIN = c(1,3), FUN = c)

```

a.) The mean is: 4.816372e+36.
```{r}
mean_18th = mean(fib_matrix1[18, ])
print(mean_18th)
```

b.) The standard deviation is: 7.582333e+15
```{r}
sd_8th = sd(fib_matrix2[, 8])
print(sd_8th)

```

c.) The entry in the 5th row of the 2nd column of the 8th matrix in fib_array is: 610
```{r}
ans = fib_matrices[5,2,8]
print(ans)
```


